\section{Introduction}

Agent-workspace and environment/cache workflows create short-lived branches,
checkpoints, cache epochs, and validation states while invoking unmodified
build and test tools. These workflows are common in agent runtimes, repository
repair systems, and environment-construction benchmarks because they need to
reuse ordinary tools while changing the state those tools should see.

The state transition becomes a filesystem problem when ordinary pathnames
continue to name objects from the wrong state. After a workspace transition, a
path bound to the previous state can cause a tool to validate the wrong
revision. A path bound to shared state can direct writes into the parent
workspace. After a cache-epoch change, a path bound to an old cache object can
feed stale inputs into a build.

The root cause is a mismatch in pathname-to-object binding or visibility for
the current state, not a need to replace lower filesystem read/write semantics.
For write-leak cases, the relevant decision is the open/create resolution that
selects the branch object or branch parent before the write occurs. After that
selection, the lower filesystem still owns write semantics.

Existing mechanisms cover this need at different ownership boundaries. FUSE can
present a dynamic view, but it makes a filesystem service own the view boundary.
Bind mounts, projected volumes, symlink forests, and OverlayFS can construct a
namespace, but the namespace must be created or updated as state changes.
Stackable filesystems can interpose beneath applications, but they own
filesystem methods and object handling~\cite{linux_fuse,kubernetes_projected_volumes,linux_overlayfs,wrapfs}.
The unresolved mismatch is the need to change only pathname-to-object bindings
or visibility while leaving ordinary filesystem ownership and semantics
unchanged.

These cases define state-transition path views, which are source-system
transitions where state changes affect only object selection or visibility. The
key insight is not that pathname lookup can choose objects. Existing
filesystems and namespace mechanisms already do that. The useful separation is
between state-dependent object selection, namespace materialization, and
filesystem-service ownership when the lower filesystem continues to own data,
writes, permissions, persistence, and consistency.

\namei realizes this separation as a policy-only object-selection boundary
inside VFS name resolution. A policy is an eBPF program attached through the
real \code{cgroup/namei_ext} path in the modified kernel. The kernel exposes one
decision function because lookup and directory enumeration share the same
contract. Given an event, pathname context, and policy state, the policy chooses
the visible lower path or visibility action while leaving the VFS object model
and lower filesystem semantics intact. Unlike a stackable or passthrough filesystem, the
policy does not implement filesystem methods, own VFS objects, run a filesystem
server, or materialize a namespace tree.

The claim is deliberately falsifiable. The evaluation begins with two upstream
transitions, one AgentFS-derived workspace transition and one
SWE-Factory-Gym-derived environment/cache transition. It classifies each
transition's oracle-relevant filesystem effects before any \namei
implementation is counted as boundary evidence. A transition that requires
workload-specific read/write/setattr interposition, synthetic contents, or
cross-path transactions is reported outside the boundary rather than counted as
\namei evidence. Same-oracle baselines then support only practical value and
cost claims.

The paper makes three planned contributions.
\begin{enumerate}
\item A state-transition path-view model that classifies source-system state
transitions into pathname-to-object selection, visibility, lower-filesystem
behavior, and out-of-model effects (Sections~\ref{sec:workload-characterization}
and~\ref{sec:design}).
\item \namei, a VFS name-resolution prototype that realizes classified
in-model effects with one eBPF decision interface for lookup and directory
enumeration, without implementing filesystem methods or owning VFS objects
(Sections~\ref{sec:design} and~\ref{sec:implementation}).
\item A correctness-first KVM evaluation plan for two examined upstream
transitions, measuring oracle correctness, operation-weighted lookup/readdir
policy events, and setup/update/runtime behavior against source-system,
distinct standalone FUSE, and materialized baselines
(Section~\ref{sec:evaluation}).
\end{enumerate}
