\section{Workload Characterization}
\label{sec:workload-characterization}

\subsection{Source Transitions}

Agent-workspace and environment/cache systems expose short-lived state
transitions that are visible through ordinary pathnames. Agent systems create
branches, copy-on-write sandboxes, and checkpoint-like workspace states.
Environment systems validate builds against cache epochs and generated
environment state.

For these transitions, choosing a broader mechanism changes more than name
resolution. A FUSE agent
filesystem can implement a dynamic view, but it also owns a filesystem service
boundary. A materialized namespace can present the correct files, but it must
construct or update that namespace as state changes. A stackable filesystem can
interpose beneath applications, but it implements filesystem methods. These are
valid mechanisms and important baselines. The question is when the upstream
transition and oracle require only policy over name resolution.

\subsection{Workload Sources}

Table~\ref{tab:transition-characterization} records the source, oracle,
policy-controlled effects, preserved lower-filesystem checks, and companion
motivation for each transition. A row supports the claim only if the chosen
transition fits the object-selection-and-visibility boundary with its own
oracle.

\begin{table}[t]
\centering
\small
\begin{tabular}{@{}p{0.15\linewidth}p{0.17\linewidth}p{0.19\linewidth}p{0.18\linewidth}p{0.17\linewidth}@{}}
\hline
Selected source & Oracle source & Policy-controlled effects & Preserved lower-FS checks & Companion motivation \\
\hline
AgentFS workspace transition & AgentFS lifecycle oracle & object selection and visibility & branch-object writes and permissions & BranchFS, Sandlock, YoloFS \\
SWE-Factory-Gym cache transition & SWE-Factory-Gym build/test oracle & cache-object selection and visibility & lower-FS reads and cache-object identity & MEnvData-SWE, SWE-rebench V2 \\
\hline
\end{tabular}
\caption{Source-derived transition characterization. Rows summarize the source and oracle fields that the KVM evaluation instantiates for the two selected transitions.}
\label{tab:transition-characterization}
\end{table}

The first selected transition is derived from AgentFS~\cite{agentfs_repo}.
BranchFS, Sandlock, and YoloFS motivate nearby agent-workspace behavior but do
not define the oracle for this row~\cite{branchfs_repo,sandlock_repo,yolofs_repo}.
The trace creates or forks a workspace state, opens a path whose parent
state also contains that name, performs a write through the workspace path,
changes to a checkpoint or sibling state, and then checks that lookup and
directory enumeration expose the expected branch object and hide stale entries.
The write itself remains a lower-filesystem write after open/create path
resolution has selected the branch object. This trace contributes evidence only
when the source oracle is expressible as lookup/readdir-time selection or
hiding of existing lower objects.

The second selected transition is an environment/cache transition from
SWE-Factory-Gym~\cite{swe_factory_repo}. MEnvData-SWE and SWE-rebench V2
motivate nearby Docker-backed environment behavior but do not define the oracle
for this row~\cite{menvagent_repo,swe_rebench_v2_repo}.
The trace begins at a cache epoch, resolves a hit to a verified local
object, resolves a miss to canonical backing, marks a local object stale or
corrupt, falls back to backing for that path, and then advances the epoch and
checks that subsequent lookup selects the new valid object. If the selected
oracle requires workload-specific file-content generation or custom write
handling, it is out of model for this paper.
