\section{Design}
\label{sec:design}

\subsection{System Overview}

\namei places a programmable decision at the point where the VFS chooses the
visible object for a pathname or directory entry. Four invariants organize the
design: (1) policy runs only at lookup and directory enumeration, (2) policy
returns bounded path-view actions over existing lower objects, (3) the kernel
validates every action before ordinary VFS execution resumes, and (4) the lower
filesystem retains permissions, file operations, data, writes, page cache,
persistence, and consistency.

Figure~\ref{fig:overview} shows the resulting flow. An application in a
workload scope performs an ordinary lookup, open, stat, exec, or readdir. The
VFS reaches the \namei decision point before committing to the visible object
for the current component or directory entry. The eBPF policy returns a bounded
path-view action, and ordinary VFS and lower-filesystem logic resumes after
kernel validation.

\begin{figure}[t]
\centering
\scriptsize
\setlength{\tabcolsep}{3pt}
\begin{tabular}{c c c c c}
\fbox{\begin{tabular}{c}workload\\cgroup\end{tabular}} &
\begin{tabular}{c}$\longrightarrow$\\lookup/readdir\end{tabular} &
\fbox{\begin{tabular}{c}VFS\\name resolution\end{tabular}} &
\begin{tabular}{c}$\longrightarrow$\\event\end{tabular} &
\fbox{\begin{tabular}{c}eBPF\\policy\end{tabular}} \\
&&&& $\downarrow$ \\
\fbox{\begin{tabular}{c}lower FS owns\\dentries, inodes, data, writes\end{tabular}} &
\begin{tabular}{c}$\longleftarrow$\\ordinary VFS path\end{tabular} &
\fbox{\begin{tabular}{c}kernel validates\\bounded action\end{tabular}} &
\begin{tabular}{c}$\longleftarrow$\\pass/redirect/hide/select\end{tabular} &
\fbox{\begin{tabular}{c}policy output\\no FS methods\end{tabular}}
\end{tabular}
\caption{\namei keeps policy at VFS name resolution. The policy chooses bounded
path-view actions, while the kernel and lower filesystem continue to own
filesystem objects and data semantics.}
\label{fig:overview}
\end{figure}

\subsection{Name-Resolution Boundary Goals}

Four goals follow from keeping policy at name resolution while leaving
filesystem ownership below the hook: expressiveness, lower-filesystem ownership,
placement cost, and fail-closed safety. Like \code{sched_ext}, \namei lets BPF
choose policy while the kernel keeps the VFS mechanisms that own objects and
execute filesystem semantics. The VFS and lower filesystem retain path walking,
dentries, inodes, permissions, file operations, page cache, writes, persistence,
and consistency.

The design defines a narrow kernel extension for name-resolution policy while
leaving full filesystem expressiveness to FUSE, stackable, or custom
filesystems. Table~\ref{tab:req-design} maps the four design goals to the RQs.

\begin{table}[t]
\centering
\scriptsize
\begin{tabular}{@{}p{0.19\linewidth}p{0.5\linewidth}p{0.17\linewidth}@{}}
\hline
Goal & Design choice & RQ \\
\hline
G1 Expressiveness & invoke one workload-scoped decision contract at lookup and directory-enumeration time & RQ1 \\
G2 Lower-filesystem ownership & return only bounded actions over existing lower objects while keeping data, writes, permissions, page cache, and persistence below the hook & RQ1, RQ3 \\
G3 Placement cost & run bounded policy at lookup/readdir instead of routing the same decision through a filesystem service & RQ2 \\
G4 Fail-closed safety & use verified eBPF outputs plus kernel validation so malformed or unsupported decisions fail visibly & RQ3 \\
\hline
\end{tabular}
\caption{Design-goal mapping. If an oracle needs synthetic contents, custom
metadata, or data-path mediation, the effect belongs outside \namei's
boundary.}
\label{tab:req-design}
\end{table}

\subsection{Policy Model}

The workloads require policy to see lookup and directory-enumeration events,
carry workload state, return only bounded path-view actions, and let the kernel
validate the result.

The policy input contains operation context, the current component, parent
context, and policy state. The policy state is workload-defined but not a
filesystem object model. For the representative workloads, policy state is a
branch or checkpoint identifier in an agent workspace, a cache or validation
epoch in an environment run, or a configuration/secret epoch in a service.

The output passes the lower lookup, redirects selection to a replacement
component, hides a name from lookup and directory enumeration, or selects a
registered lower object. Access-control denial remains outside the model
because \namei focuses on selecting or hiding existing objects during name
resolution. Directory enumeration uses the same decision function so that
aliases returned by readdir match the objects selected by lookup.

The kernel checks the decision, preserves normal permission and
lower-filesystem checks for the selected object, maintains ordinary VFS path
resolution invariants, and fails invalid decisions visibly instead of falling
back to a weaker policy.

\subsection{Filesystem Ownership Boundary}

\namei excludes filesystem ownership by leaving inode operations, file
operations, dentry and inode allocation, synthetic contents, custom metadata,
read/write mediation, distributed indexes, and cross-path transactions outside
the hook.

The name-resolution boundary helps only when the source transition can
be expressed as lookup selection or directory visibility. When the oracle
requires data-path mediation, synthetic content, custom metadata, or
application-level orchestration, the transition remains motivation or related
work rather than \namei evidence.
