\section{Evaluation}
\label{sec:evaluation}

\subsection{Research Questions}

The evaluation is organized around three research questions. Each answer is
conditioned on a source oracle: the same workload behavior must be used for
\namei, the feature-equivalent FUSE row, and the custom/stackable
responsibility comparison.
One-off mechanism checks, source surveys, and weak baselines do not answer an
RQ by themselves.

\begin{description}
\item[RQ1] Can a narrow VFS name-resolution extension express real
state-dependent path-view policies without taking over filesystem semantics?
\item[RQ2] What is the cost of putting programmable policy on the VFS
name-resolution path compared with a feature-equivalent FUSE policy
implementation?
\item[RQ3] Does \namei provide a narrower verifier-bounded, fail-closed
ownership boundary than building a custom or stackable filesystem when the
needed policy is only name resolution?
\end{description}

\subsection{Experimental Setup}

All \namei mechanism runs execute in KVM~\cite{linux_kvm} with the modified
kernel and the real \code{cgroup/namei_ext} attach path. Make targets own the
run, attach the policy, collect raw artifacts, and fail on missing capability,
verifier failure, syscall failure, or oracle failure. Each final row records the
kernel commit, image identity, policy object, workload inputs, stdout/stderr,
dmesg, per-operation lookup/readdir events, and comparison-specific setup
observations.
Operation-weighted metrics weight each policy action by its observed frequency
in the workload trace.

Each final reported run records the hardware and kernel configuration, KVM
image identity, filesystem configuration, run count, warmup rule,
confidence-interval method, cache-hot/cache-cold protocol, feature-equivalent
FUSE implementation, and the custom or stackable filesystem boundary account for
the workload.

Correctness gates every number. A FUSE row is used for RQ2 only after it
implements the same policy over the same source oracle. No-hook and
lower-filesystem rows calibrate fixed hook and measurement cost. They are not
competing baselines. RQ3 uses the same oracle to compare ownership and
containment rather than to claim that a custom/stackable filesystem was
reimplemented as an executed baseline.

\subsection{RQ1: Expressiveness And Sufficiency}

RQ1 asks whether the VFS boundary can express path-view transitions extracted
from source oracles while preserving lower-filesystem semantics. A reportable
row must pass the source oracle through the real KVM attach path and show that
\namei changed only lookup or directory visibility, while runtime orchestration,
storage state, synthetic contents, writes, and data-path behavior remain with
the source system or lower filesystem.

\begin{table}[t]
\centering
\scriptsize
\begin{tabular}{@{}L{0.17\linewidth}L{0.27\linewidth}L{0.25\linewidth}L{0.21\linewidth}@{}}
\hline
Workload & Source oracle & \namei evidence & Result evidence \\
\hline
Agent workspace lifecycle~\cite{agentfs_repo} & AgentFS-derived branch or
checkpoint state selects or hides branch-visible objects. The oracle checks
final tree state, whiteout/hide effects, symlink metadata, lookup/readdir
coherence, and branch-visible state. & KVM \code{cgroup/namei_ext} policy
selects or hides existing lower objects while the lower filesystem keeps
permissions, writes, symlink metadata, file data, and ordinary directory
operations. & 3/3 KVM runs passed the fixed trace and final-tree oracle with
zero failed rows. Each run observed 2,581 lookups, 903 readdir decisions, 419
selections, and both lookup/readdir hides; direct checks placed create, rename,
and unlink mutations only in the selected lower tree. \\
Traditional build/cache transition~\cite{swe_factory_repo,menvagent_repo,swe_rebench_v2_repo} &
Cache epoch selects a verified local object, canonical fallback, or hides
stale/corrupt local state. The oracle checks source build/test success and
stale/corrupt non-use. & \namei selects existing verified local or canonical
objects while cache population, content validation, build/test execution, and
writes stay outside the hook. & \textbf{[Result]} cache-state matrix covering
hit, miss, stale, corrupt, and epoch update states. \\
\hline
\end{tabular}
\caption{RQ1 answer table. A row becomes evidence only after the source-oracle
run passes through the real KVM attach path and lower-filesystem semantics are
preserved.}
\label{tab:rq1-evidence}
\end{table}

The required RQ1 evidence is a reviewed KVM pass/fail matrix that preserves
lookup/readdir coherence and lower-filesystem behavior for each admitted source
oracle. A workload is outside \namei's boundary if the oracle needs synthetic
contents, data-path mediation, write-conflict handling, or custom metadata
ownership in the common path.

\subsection{RQ2: Cost Versus FUSE}

RQ2 isolates the placement cost of policy by comparing \namei with a
feature-equivalent FUSE policy service under the same oracle. Non-name-resolution
lifecycle behavior is delegated identically in both rows. RQ2 does not compare
different runtimes. Instead, it compares placing the same lookup/readdir
decision inside VFS name resolution with placing it in a FUSE request path.

\begin{table}[t]
\centering
\scriptsize
\begin{tabular}{@{}L{0.2\linewidth}L{0.26\linewidth}L{0.23\linewidth}L{0.21\linewidth}@{}}
\hline
Workload & Feature-equivalent FUSE comparison & Metrics & Result evidence \\
\hline
Agent workspace lifecycle & \namei policy in KVM versus a FUSE policy service
that implements the same source oracle. & lookup, open, stat, access, exec, and
readdir latency; macro lifecycle runtime; action mix; FUSE request-path cost &
\textbf{[Result]} correctness-gated latency and macro-runtime table with
confidence intervals and FUSE option record. \\
Traditional build/cache transition & \namei cache policy in KVM versus a FUSE
cache-policy service that implements the same cache-state oracle. & cache-hot
and cache-cold lookup, open, stat, access, and readdir latency; build/test macro
runtime; operation-weighted invocation rate & \textbf{[Result]}
correctness-gated cache-state overhead table. \\
\hline
\end{tabular}
\caption{RQ2 evidence table. FUSE numbers are interpreted only after the FUSE row
implements the same source policy and passes the same source oracle.}
\label{tab:rq2-evidence}
\end{table}

RQ2 interprets only correctness-gated rows with feature-equivalent FUSE
implementations. The result compares lookup/readdir placement cost while keeping
correctness and delegated runtime behavior fixed. A workload without a
feature-equivalent FUSE row is not interpreted for RQ2.
The required RQ2 evidence is a correctness-gated latency and macro-runtime table
for the admitted source oracles, including pass-through overhead,
action-specific overhead, cache-hot/cache-cold behavior, operation-weighted
invocation rate, matched action mix, and FUSE daemon/request-path cost.
Calibration rows are not treated as competing baselines.

\subsection{RQ3: Ownership And Containment Boundary}

RQ3 compares ownership boundaries for the same oracle-relevant behavior. The
question is not whether custom or stackable filesystems are expressive, because
they are. The question is whether the workload needs that broader boundary when
the policy effect is only name-resolution selection or visibility.

\begin{table}[t]
\centering
\scriptsize
\setlength{\tabcolsep}{2pt}
\begin{tabular}{@{}L{0.14\linewidth}L{0.15\linewidth}L{0.16\linewidth}L{0.16\linewidth}L{0.17\linewidth}L{0.14\linewidth}@{}}
\hline
Source behavior & Oracle operation & \namei responsibility & Kernel/lower-FS responsibility & Custom/stackable responsibility & Fail-closed evidence \\
\hline
Agent workspace branch/checkpoint state & lookup and readdir select or hide
branch-visible objects & eBPF returns bounded selection/visibility decisions &
kernel validates targets and malformed outputs; lower FS owns data, writes,
permissions, symlink metadata, and page-cache behavior & custom/stackable workspace FS would own or mediate
lookup, readdir, create, unlink, rename, COW/checkpoint state, daemon state, and
data-path behavior & \textbf{[Result]} invalid action, bad target, and
malformed-name rejection on the real attach path \\
Build/cache cache-state transition & lookup/open selects verified local
or canonical objects, or hides stale/corrupt state & eBPF returns bounded object
selection or visibility decisions & source evaluator owns cache population and
validation; lower FS owns writes, data path, persistence, and page-cache behavior & custom/cache FS
would own or mediate cache metadata, stale/corrupt state, open/read/write behavior, daemon
state, and persistence decisions & \textbf{[Result]} invalid action, bad target,
and malformed-name rejection on the real attach path \\
\hline
\end{tabular}
\caption{RQ3 boundary table. The comparison is responsibility and containment for
the same oracle-relevant behavior, not a claim that custom filesystems are less
expressive.}
\label{tab:rq3-evidence}
\end{table}

RQ3 is supported when the same-oracle rows pass, malformed decisions fail
closed, and the responsibility comparison shows that a broader filesystem
mechanism would own methods, daemon state, storage semantics, or data-path
behavior that \namei leaves to VFS, the lower filesystem, or the source runtime.
The required RQ3 evidence is a workload-specific ownership table tied to the
admitted source oracles plus invalid-policy KVM rows on the real
\code{cgroup/namei_ext} path. Per-source ownership rows must name owned
filesystem methods, state responsibility, privileged code surface, and policy
code size with source evidence.

\subsection{Evaluation Scope}

The main evaluation has two primary workload families, reported only after
reviewed same-oracle KVM runs: Agent workspace and traditional build/cache.
Build/environment datasets may supply build/test oracles, but the second
workload is not an agent workload. A row is interpreted only after the same
source oracle passes for the compared mechanisms. Service/config remains
motivating scope until a concrete source oracle depends on lookup-time object
selection. Materialized
namespace mechanisms, eBPF LSM hooks, and native production mechanisms remain
related work and background unless a selected source workload makes one of them
part of the source behavior being evaluated.
This draft has completed the reviewed Agent-workspace RQ1 row for the fixed
AgentFS-derived path-view slice. The independent traditional build/cache row,
the correctness-gated \namei/FUSE latency and macro-runtime rows, and the
workload-specific ownership and invalid-policy rows remain incomplete; the
paper therefore does not yet claim complete answers to RQ1--RQ3.
