\section{Related Work}
\label{sec:related-work}

Related work falls along a spectrum of filesystem ownership. Namespace
mechanisms construct views before lookup, and access-control hooks mediate or
observe operations near the VFS. \namei puts bounded object-selection policy in
VFS name resolution, while FUSE, stackable filesystems, custom filesystems, and
metadata services require broader filesystem-service or object-model ownership.

\subsection{FUSE Policy Services}

FUSE~\cite{linux_fuse}, FUSE passthrough~\cite{linux_fuse_passthrough},
ExtFUSE~\cite{extfuse}, and FUSE-BPF~\cite{fuse_bpf_patch} demonstrate the
value of placing filesystem policy in, or accelerating, a filesystem-service
boundary. RFUSE, CoFS, and DFUSE further show that this boundary can be
optimized or specialized with scalable kernel-userspace communication, fixed
container-image lookup structures, and distributed-cache consistency
support~\cite{rfuse_fast24,cofs_fast26,distfuse_arxiv25}. These systems make
FUSE the RQ2 baseline family, not a strawman: a fair comparison must implement
the same path-view policy under the same source oracle and justify the FUSE
caching/passthrough configuration. \namei instead keeps policy needed by the
source oracle at VFS name resolution when the workload only needs path
selection or visibility over existing objects.

\subsection{Custom, Stackable, And Metadata-Service Filesystems}

Wrapfs~\cite{wrapfs}, Bento~\cite{bento}, and custom filesystems demonstrate
broader filesystem-extension boundaries. They can implement richer behavior and,
in Bento's case, reduce the risk of in-kernel filesystem development. They are
the RQ3 ownership comparison point because the developer still owns filesystem
methods, storage semantics, or framework-specific filesystem behavior. These
mechanisms remain the appropriate implementation boundary when the workload
needs custom file operations, synthetic contents, persistence, or richer
semantics.

DeltaFS, IndexFS, and TableFS show that full metadata services and specialized
filesystems are valuable when the problem is metadata distribution, indexing,
or storage layout~\cite{deltafs_sc21,indexfs,tablefs}. Recent systems reinforce
the same boundary: Oxbow coordinates kernel, userspace, and device components;
DeLFS builds a decentralized kernel filesystem; FalconFS moves path resolution
and metadata indexing into a distributed filesystem service; SwitchFS and
MesaFS target distributed metadata updates; and SYSSPEC/SpecFS generates a
complete filesystem from specifications~\cite{oxbow_osdi26,delfs_osdi26,falconfs_nsdi26,switchfs_eurosys26,mesafs_eurosys26,specfs_fast26}.
They are not the main workload baselines for \namei, but they clarify the
boundary because such workloads require a metadata service or full filesystem
when the oracle needs distributed metadata, data-path ownership, persistence,
or a new filesystem object model.

\subsection{Namespace Construction}

Bind mounts, projected volumes, symlink forests, copy trees, and OverlayFS are
practical ways to present alternate views~\cite{kubernetes_projected_volumes,linux_overlayfs}.
They preserve ordinary kernel filesystem behavior, but express policy by
constructing or updating namespace state. \namei asks when the policy can stay
at lookup and directory-enumeration time instead of materializing a view before
the transition completes.

\subsection{Access Control And Observation Hooks}

Landlock, BPF LSM programs, and fanotify show that Linux already exposes
specialized hooks around filesystem access and events~\cite{linux_landlock,linux_bpf_lsm,linux_fanotify}.
Among neighboring mechanisms, these hooks sit between materialized namespace
construction and \namei. They can run verified policy in the kernel, but their
natural role is to deny, mediate, or observe operations. \namei selects which
existing object a name denotes during resolution and then returns control to the
ordinary VFS and lower filesystem. Access-control denial remains outside the
current action set because the primary claim is name-resolution selection and
visibility. USEC is a recent access-control neighbor with an LSM-compatible
mandatory-access-control design~\cite{usec_osdi26}; XRP, vBPF, PeeR,
KRAKENGUARD, Xkernel, and bpftime/EIM are adjacent programmability work on
storage-function hooks, eBPF virtualization, eBPF scheduling, eBPF isolation,
kernel tunability, and userspace extension safety~\cite{xrp_osdi22,vbpf_osdi26,peer_osdi26,krakenguard_nsdi26,xkernel_osdi26,bpftime_eim_osdi25}.
They help set safety and overhead expectations for programmable kernel
extensions, but they do not provide a VFS name-resolution object-selection
boundary.

\subsection{Source Workload Systems}

Agent filesystems are closest to our workloads, but they are workload sources
rather than direct cost baselines~\cite{agentfs_repo,branchfs_repo,sandlock_repo,yolofs_repo}.
They show that agent workspaces need branch, sandbox, staging, and
snapshot-like views, while their implementations also own behavior beyond
\namei's boundary.
For agent workloads, custom or stackable filesystem ownership provides the RQ3
responsibility comparison, while FUSE remains the RQ2 cost comparison. \namei applies when the
same source oracle can be satisfied by changing path visibility or object
selection while leaving the rest of that ownership with the source system or
lower filesystem.

Environment and cache systems provide a second workload source, not a filesystem
mechanism baseline. SWE-Factory-Gym, MEnvData-SWE, and SWE-rebench V2 provide
build/test oracles for environment state and cache reuse
policies~\cite{swe_factory_repo,menvagent_repo,swe_rebench_v2_repo}. CoFS and
large-scale LLM data-pipeline work add contemporary evidence that build,
container, and data-cache paths matter to real systems~\cite{cofs_fast26,llm_data_pipeline_osdi26}.
Agentic systems such as Murakkab and SREGym show that modern agents run through
multi-step tool, workflow, and live-environment oracles~\cite{murakkab_osdi26,sregym_arxiv26}.
These sources motivate workload state and oracle selection. Their runtime
orchestration, environment synthesis, projected-volume materialization, and
reload behavior remain outside \namei's mechanism claim.
