\documentclass[10pt]{article}

\usepackage[UTF8]{ctex}
\usepackage[margin=1in]{geometry}
\usepackage{array}
\usepackage{hyperref}
\usepackage{xspace}
\setlength{\emergencystretch}{3em}
\AtBeginDocument{\sloppy}

\newcommand{\namei}{\textsc{namei\_ext}\xspace}
\newcommand{\code}[1]{\texttt{\detokenize{#1}}}
\newcommand{\file}[1]{\path{#1}}
\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}

\title{\namei: A \code{sched_ext}-Style Extension Point for VFS Name Resolution}
\author{Anonymous}
\date{}

\begin{document}
\maketitle

\begin{abstract}
Modern agent, build, and service systems increasingly construct task-specific
filesystem views around ordinary tools. The risk is binding a pathname to the
wrong lower object, which can corrupt an agent workspace, reuse stale build
state, or expose an unintended service configuration. The root cause is a
boundary mismatch in which workload state changes pathname-to-object selection
during name resolution, while permissions, data operations, writes,
persistence, and consistency should remain lower-filesystem responsibilities.
Existing mechanisms either materialize a namespace before lookup, mediate access
through security hooks, or cross into broader FUSE, stackable, or custom
filesystem ownership. We argue that the missing narrow boundary is policy at VFS
name resolution, where pathname-to-object selection and directory visibility are
decided. Realizing this boundary is hard because policy must run at lookup and
readdir, return bounded kernel-validated actions, and remain comparable under the
same correctness condition against FUSE and broader filesystem boundaries.
\namei implements this ownership split as a \code{sched_ext}-style VFS extension point.
One \code{cgroup/namei_ext} decision function lets eBPF choose pass, redirect,
hide, and \code{SELECT_TARGET} actions over kernel-registered lower paths while
the kernel retains filesystem ownership. The evaluation uses source-derived
agent/workspace and environment/cache correctness conditions to compare \namei
with feature-equivalent FUSE and to identify which filesystem responsibilities
remain outside the hook. Service/config remains conditional on an admitted
lookup-time oracle.
\end{abstract}

\input{sections/01-introduction}
\input{sections/02-background}
\input{sections/02-motivation}
\input{sections/03-design}
\input{sections/04-implementation}
\input{sections/05-evaluation}
\input{sections/07-limitations}
\input{sections/06-related-work}
\input{sections/08-conclusion}

{\small
\bibliographystyle{plain}
\bibliography{refs}
}

\end{document}
