\documentclass[10pt]{article}

\usepackage[UTF8]{ctex}
\usepackage[margin=1in]{geometry}
\usepackage{array}
\usepackage{hyperref}
\usepackage{xspace}
\setlength{\emergencystretch}{3em}
\AtBeginDocument{\sloppy}

\newcommand{\namei}{\textsc{namei\_ext}\xspace}
\newcommand{\code}[1]{\texttt{\detokenize{#1}}}
\newcommand{\file}[1]{\path{#1}}
\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}

\title{\namei: A \code{sched_ext}-Style Extension Point for VFS Name Resolution}
\author{Anonymous}
\date{}

\begin{document}
\maketitle

\begin{abstract}
Modern agent, build, and service systems increasingly construct task-specific
filesystem views around ordinary tools. The risk is binding a pathname to the
wrong lower object, which can corrupt an agent workspace, reuse stale build
state, or expose an unintended service configuration. The root cause is a
boundary mismatch in which workload state changes pathname-to-object selection
during name resolution, while permissions, data operations, writes,
persistence, and consistency should remain lower-filesystem responsibilities.
Existing
mechanisms either materialize a namespace before lookup, mediate access through
security hooks, or cross into broader FUSE, stackable, or custom filesystem
ownership. We argue that the missing narrow boundary is policy at VFS name
resolution, where pathname-to-object selection and directory visibility are
decided. Realizing this boundary is hard because policy must run at lookup and
readdir, return bounded kernel-validated actions, and remain comparable under
the same correctness condition against FUSE and broader filesystem boundaries.
\namei explores this boundary as a \code{sched_ext}-style VFS extension point in which
eBPF chooses bounded lookup and directory-enumeration policy while the kernel
retains filesystem ownership. The prototype implements one decision function,
the \code{cgroup/namei_ext} attachment, and kernel validation for pass,
redirect, hide, and registered directory-target selection. The evaluation uses
correctness oracles from agent/workspace and environment/cache systems, with
environment/cache file-object evidence gated on final-file target selection, to
ask whether this narrow boundary can satisfy real object-selection and
directory-visibility policies, avoid FUSE request-path cost, and provide a
narrower verifier-bounded, fail-closed ownership boundary than custom or
stackable filesystems.
\end{abstract}

\input{sections/01-introduction}
\input{sections/02-background}
\input{sections/02-motivation}
\input{sections/03-design}
\input{sections/04-implementation}
\input{sections/05-evaluation}
\input{sections/07-limitations}
\input{sections/06-related-work}
\input{sections/08-conclusion}

{\small
\bibliographystyle{plain}
\bibliography{refs}
}

\end{document}
